---
title: "Project_1"
author: "ZZ"
date: "12/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

What is mean total number of steps taken per day?

1. Calculation of total number of step taken per day.
```{r}
library(ggplot2)
library(dplyr)
setwd("~/Documents/jhdatascience")
activity <- read.csv("activity.csv")

activity$date <- as.Date(activity$date)

stepsPerDay <- activity %>%
        group_by(date) %>%
        summarize(sumsteps = sum(steps, na.rm = TRUE)) 
head(stepsPerDay,10)
```
2. a histogram of the total number of steps taken each day.
```{r}
hist(stepsPerDay$sumsteps, main = "Histogram of Daily Steps", 
     col="red", xlab="Steps", ylim = c(0,30))
```
3. Calculate and report the mean and median of the total number of steps taken per day.
```{r}
meanPreNA <- round(mean(stepsPerDay$sumsteps),digits = 2)
medianPreNA <- round(median(stepsPerDay$sumsteps),digits = 2)

print(paste("The mean is: ", meanPreNA))
print(paste("The median is: ", medianPreNA))

```

What is the average daily activity pattern?
1. step by time interval
```{r}
stepsPerInterval <- activity %>%
        group_by(interval) %>%
        summarize(meansteps = mean(steps, na.rm = TRUE)) 
#Display first 10 rows of data
head(stepsPerInterval,10)
plot(stepsPerInterval$meansteps ~ stepsPerInterval$interval,
     col="black", type="l", xlab = "5 Minute Intervals", ylab = "Average Number of Steps",
     main = "Steps By Time Interval")
```
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
print(paste("Interval containing the most steps on average: ",stepsPerInterval$interval[which.max(stepsPerInterval$meansteps)]))

print(paste("Average steps for that interval: ",round(max(stepsPerInterval$meansteps),digits=2)))
```
Imputing missing value
   1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
print(paste("The total number of rows with NA is: ", sum(is.na(activity$steps))))
```
2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
Answer: There are several strategy to do this, one way that is listed in the example is the midrange values, which take the mean and median numuric values, it is easy to compare, the disvantage of this method would bethe posibility of overfitting.  The other way would be imputation, specifically perturbation.

3.Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
head(activity,15)
activityNoNA <- activity  
for (i in 1:nrow(activity)){
        if(is.na(activity$steps[i])){
                activityNoNA$steps[i]<- stepsPerInterval$meansteps[activityNoNA$interval[i] == stepsPerInterval$interval]
        }
}
head(activityNoNA,15)
```
4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}

stepsPerDay <- activityNoNA %>%
        group_by(date) %>%
        summarize(sumsteps = sum(steps, na.rm = TRUE)) 
head(stepsPerDay,5)

hist(stepsPerDay$sumsteps, main = "Histogram of Daily Steps", 
     col="black", xlab="Steps")

meanPostNA <- round(mean(stepsPerDay$sumsteps), digits = 2)
medianPostNA <- round(median(stepsPerDay$sumsteps), digits = 2)

print(paste("mean = ", mean(meanPostNA)))
print(paste("median = ", median(medianPostNA)))

NACompare <- data.frame(mean = c(meanPreNA,meanPostNA),median = c(medianPreNA,medianPostNA))
rownames(NACompare) <- c("Pre NA Transformation", "Post NA Transformation")
print(NACompare)

```

Are there differences in activity patterns between weekdays and weekends?

```{r}

activityDoW <- activityNoNA
activityDoW$date <- as.Date(activityDoW$date)
activityDoW$day <- ifelse(weekdays(activityDoW$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")
activityDoW$day <- as.factor(activityDoW$day)

activityWeekday <- filter(activityDoW, activityDoW$day == "weekday")
activityWeekend <- filter(activityDoW, activityDoW$day == "weekend")

activityWeekday <- activityWeekday %>%
        group_by(interval) %>%
        summarize(steps = mean(steps)) 
activityWeekday$day <- "weekday"

activityWeekend <- activityWeekend %>%
        group_by(interval) %>%
        summarize(steps = mean(steps)) 
activityWeekend$day <- "weekend"

wkdayWkend <- rbind(activityWeekday, activityWeekend)
wkdayWkend$day <- as.factor(wkdayWkend$day)


g <- ggplot (wkdayWkend, aes (interval, steps))
g + geom_line() + facet_grid (day~.) + 
        theme(axis.text = element_text(size = 12),axis.title = element_text(size = 14)) + 
        labs(y = "Number of Steps") + labs(x = "Interval") + 
        ggtitle("Average Number of Steps - Weekday vs. Weekend") + 
        theme(plot.title = element_text(hjust = 1.0))

```


The graph demonstrate differences in the patterns on the average daily intervals. 
For weekdays - there are larger spike in the morning intervals which could indicate their activites of morning work out or commuting to work. Fod the weekend- the acitivites are more balances as the individual may doing all kind of activites instead of working or sitting in an set hours time of time.